ML model comparison

### 2020-10-03 src/s_ml_comparison.Rmd
Update s_ml_comparison.Rmd

f_replicate_ml_variety <- function(x, method_, caretGrid_=NULL){
  temp <- x
  a_array <- array(dim=c(2000, 4, 1))
  a_array[,,1] <- temp
  a_accuracy <- r_star_ml_variety(a_array, method_=method_, caretGrid = caretGrid_)
  return(a_accuracy)
}

@@

tunegrid_rf <- tibble(mtry = 1:2)
tunegrid_gbm <- expand.grid(interaction.depth=c(3), 
                             n.trees = 50,
                             shrinkage=c(0.1),
                             n.minobsinnode=10)
tunegrid_knn <- tibble(k = c(5, 10, 15, 20, 40))
tunegrid_svm <- tibble(C = c(0.25, 0.5, 0.75))
tunegrid_nnet <- tibble(size=c(2, 3, 4, 5), decay=c(0.5, 0.5, 0.5, 0.5))
niterates <- 50
m_res <- matrix(nrow=niterates, ncol = 6)
l_rf_hyper <- vector(length = niterates, mode="list")
l_knn_hyper <- vector(length = niterates, mode="list")
l_svm_hyper <- vector(length = niterates, mode="list")
l_nnet_hyper <- vector(length = niterates, mode="list")
for(i in 1:niterates){
  print(i)
  x <- f_generate_lower_var_four(1/3, 0.3, 1, 2000)
  m_res[i, 1] <- f_replicate_ml_variety(x, "gbm", tunegrid_gbm)
  m_res[i, 2] <- f_replicate_rhat(x)
  temp <- f_replicate_ml_variety_gridded(x, method_="rf", caretGrids_ = tunegrid_rf)
  l_rf_hyper[[i]] <- temp$hyper
  m_res[i, 3] <- temp$a_accuracy
  temp <- f_replicate_ml_variety_gridded(x, "knn", tunegrid_knn)
  l_knn_hyper[[i]] <- temp$hyper
  m_res[i, 4] <- temp$a_accuracy
  temp <- f_replicate_ml_variety_gridded(x, "svmLinear", tunegrid_svm)
  l_svm_hyper[[i]] <- temp$hyper
  m_res[i, 5] <- temp$a_accuracy
  temp <- f_replicate_ml_variety_gridded(x, "nnet", tunegrid_nnet)
  l_nnet_hyper[[i]] <- temp$hyper
  m_res[i, 6] <- temp$a_accuracy
}

@@

fit_cp <- sampling(
  model_cp, data = eight_schools,
  iter = 4000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)

fit_ncp <- sampling(
  model_ncp, data = eight_schools,
  iter = 4000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)

@@

>>>>

f_replicate_ml_variety <- function(x, method_, caretGrid_=NULL){
  temp <- x
  a_array <- array(dim=c(nrow(x), 4, 1))
  a_array[,,1] <- temp
  a_accuracy <- r_star_ml_variety(a_array, method_=method_, caretGrid = caretGrid_)
  return(a_accuracy)
}

@@

tunegrid_rf <- tibble(mtry = 1:2)
tunegrid_gbm <- expand.grid(interaction.depth=c(3), 
                             n.trees = 50,
                             shrinkage=c(0.1),
                             n.minobsinnode=10)
tunegrid_knn <- tibble(k = c(5, 10, 15, 20, 40))
tunegrid_svm <- tibble(C = c(0.25, 0.5, 0.75))
tunegrid_nnet <- tibble(size=c(2, 3, 4, 5), decay=c(0.5, 0.5, 0.5, 0.5))
tunegrid_multinom <- tibble(decay=c(0.1, 0.2, 0.5, 1))
niterates <- 50
m_res <- matrix(nrow=niterates, ncol = 7)
l_rf_hyper <- vector(length = niterates, mode="list")
l_knn_hyper <- vector(length = niterates, mode="list")
l_svm_hyper <- vector(length = niterates, mode="list")
l_nnet_hyper <- vector(length = niterates, mode="list")
l_multinom_hyper <- vector(length = niterates, mode="list")
for(i in 1:niterates){
  print(i)
  x <- f_generate_lower_var_four(1/3, 0.3, 1, 2000)
  m_res[i, 1] <- f_replicate_ml_variety(x, "gbm", tunegrid_gbm)
  m_res[i, 2] <- f_replicate_rhat(x)
  temp <- f_replicate_ml_variety_gridded(x, method_="rf", caretGrids_ = tunegrid_rf)
  l_rf_hyper[[i]] <- temp$hyper
  m_res[i, 3] <- temp$a_accuracy
  temp <- f_replicate_ml_variety_gridded(x, "knn", tunegrid_knn)
  l_knn_hyper[[i]] <- temp$hyper
  m_res[i, 4] <- temp$a_accuracy
  temp <- f_replicate_ml_variety_gridded(x, "svmLinear", tunegrid_svm)
  l_svm_hyper[[i]] <- temp$hyper
  m_res[i, 5] <- temp$a_accuracy
  temp <- f_replicate_ml_variety_gridded(x, "nnet", tunegrid_nnet)
  l_nnet_hyper[[i]] <- temp$hyper
  m_res[i, 6] <- temp$a_accuracy
  temp <- f_replicate_ml_variety_gridded(x, "multinom", tunegrid_multinom)
  l_nnet_hyper[[i]] <- temp$hyper
  m_res[i, 7] <- temp$a_accuracy
}

@@

fit_cp <- sampling(
  model_cp, data = eight_schools,
  iter = 2000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)

fit_ncp <- sampling(
  model_ncp, data = eight_schools,
  iter = 2000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)

@@

Looking into rf performance on 8 schools model more closely
```{r}
x <- x_cp
x <- split_data(x)

nparams <- dim(x)[3]
nchains <- dim(x)[2]
niter <- dim(x)[1]
m_flattened <- matrix(nrow = nrow(x) * nchains,
                      ncol = (nparams + 1))
k <- 1
for(i in 1:nchains){
  for(j in 1:nrow(x)){
    m_flattened[k, 1:nparams] <- x[j, i, ]
    m_flattened[k, (nparams + 1)] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame()
colnames(m_flattened)[nparams + 1] <- "chain"
r <- m_flattened %>% 
  mutate(chain=as.factor(chain))

rand_samples <- sample(1:nrow(r), 0.7 * nrow(r))
training_data <- r[rand_samples, ]
testing_data <- r[-rand_samples, ]
gbmFit <- train(chain ~ ., data = training_data, 
               trControl = trainControl(method = 'none'),
               method="rf",
               tuneGrid = tibble(mtry=1))
plda <- predict(object=gbmFit, newdata=testing_data)
a_accuracy <- 
  tibble(predicted=plda, actual=testing_data$chain) %>%
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct)) %>% 
  pull()
a_accuracy * n_distinct(training_data$chain)
varImp(gbmFit)

x[ , , 9] %>%
  as.data.frame() %>%
  mutate(iter=seq(1,1000,1)) %>%
  melt(id.vars="iter") %>%
  ggplot(aes(x=value, fill=as.factor(variable))) +
  stat_density(position="identity", alpha=0.2)

tibble(var1=testing_data$V1, var2=testing_data$V9, chain=plda) %>% 
  ggplot(aes(x=var1, y=var2, colour=as.factor(chain))) +
  geom_jitter()
```

Cauchy model
```{r}
fit_nom <- stan(file = 'cauchy_nom.stan', seed = 7878,
                refresh = 0)
x <- rstan::extract(fit_nom, permuted=F)
temp1 <- f_replicate_ml_simple(x, "gbm", tunegrid_gbm)
temp <- f_replicate_ml_simple_gridded(x, "rf", tibble(mtry=1))
temp$a_accuracy

fit_alt1 <- stan(file = 'cauchy_alt_1.stan',
                 seed = 7878, refresh = 0)
x <- rstan::extract(fit_alt1, permuted=F)
temp1 <- f_replicate_ml_simple(x, "gbm", tunegrid_gbm)
temp <- f_replicate_ml_simple_gridded(x, "rf", tibble(mtry=1))
temp$a_accuracy
```

AR1 vs independent
```{r}
f_generate_lower_var_four_independent <- function(var_ratio, rho, sigma, L){
  x <- matrix(nrow = L, ncol = 4)
  for(i in 1:3)
    x[, i] <- rnorm(L, 0, sqrt(sigma^2/(1-rho^2)))
  z <- rnorm(L, 0, sqrt(var_ratio * sigma^2/(1-rho^2)))
  x[, 4] <- z
  return(x)
}

rho <- 0.95
lgbm <- vector(length=2)
lrf <- vector(length=2)
x <- f_generate_lower_var_four(1/3, rho, 1, 10000)
var(x[, 4])
lgbm[1] <- f_replicate_ml_variety(x, "gbm", tunegrid_gbm)
lrf[1] <- f_replicate_ml_variety(x, "rf", tibble(mtry=1))

x <- f_generate_lower_var_four_independent(1/3, rho, 1, 10000)
var(x[, 4])
lgbm[2] <- f_replicate_ml_variety(x, "gbm", tunegrid_gbm)
lrf[2] <- f_replicate_ml_variety(x, "rf", tibble(mtry=1))
```

Try optimal Bayes classifier: two normals
```{r}
L <- 10000
x1 <- rnorm(L, 0, 1)
x2 <- rnorm(L, 1, 1)
x <- matrix(c(x1, x2), byrow = F, ncol=2)

a_array <- array(dim=c(nrow(x), 2, 1))
a_array[,,1] <- x
r_star_ml_variety(a_array, method_ ="rf", caretGrid = tibble(mtry=2))
```

Optimal classifier any dimension data
```{r}
# generate a correlation matrix via LKJ
rlkj <- function(n, nu, d){
  r_list <- vector(length = n, mode = 'list')
  for(i in 1:n){
    if(d==1)
      r = as.array(1)
    else if(d==2){
      rho <- 2 * rbeta(1, nu, nu) - 1
      r <- matrix(c(1, rho, rho, 1), ncol = 2)
    }else{
      beta <- nu + (d - 2) / 2
      u <- rbeta(1, beta, beta)
      r_12 <- 2 * u - 1
      r <- matrix(c(1, r_12, r_12, 1), ncol = 2)
      for(m in 2:(d - 1)){
        beta <- beta - 0.5
        y <- rbeta(1, m / 2, beta)
        a <- rnorm(m)
        anorm <- sqrt(sum(a^2))
        u <- a / anorm
        w <- sqrt(y) * u
        A <- chol(r)
        z <- w%*%A
        r <-cbind(r, t(z))
        r <- rbind(r, c(z, 1))
      }
    }
    r_list[[i]] <- r
  }
  return(r_list)
}

nd <- 50
df <- 200
Sigma1 <- rlkj(1, 50, nd)[[1]]
Sigma2 <- rlkj(1, 50, nd)[[1]]

# generate any dimensional data
library(mvtnorm)
n <- 2000
x1 <- rmvnorm(n, rep(0, nd), Sigma1)
x2 <- rmvnorm(n, rep(0, nd), Sigma2)

# calculate likelihood from each mode for each point
lp11 <- map_dbl(seq(1, nrow(x1), 1), ~dmvnorm(x1[., ], rep(0, nd), Sigma1, log = T))
lp12 <- map_dbl(seq(1, nrow(x1), 1), ~dmvnorm(x1[., ], rep(0, nd), Sigma2, log = T))
lp21 <- map_dbl(seq(1, nrow(x2), 1), ~dmvnorm(x2[., ], rep(0, nd), Sigma1, log = T))
lp22 <- map_dbl(seq(1, nrow(x2), 1), ~dmvnorm(x2[., ], rep(0, nd), Sigma2, log = T))
r_star_opt <- 0.5 * (mean(lp11>lp12) + mean(lp22>lp21)) / 0.5

# calculate R* from gbm and rf
x <- array(dim=c(nrow(x1), 2, ncol(x1)))
x[, 1, ] <- x1
x[, 2, ] <- x2
r_star(x, split_chains = T)

f_replicate_ml_variety_simple <- function(x, method_, caretGrid_=NULL){
  temp <- x
  a_accuracy <- r_star_ml_variety(x, method_=method_, caretGrid = caretGrid_)
  return(a_accuracy)
}
f_replicate_ml_variety_simple(x, "rf", tibble(mtry=20))
f_replicate_ml_variety_simple(x, "gbm",
                              tibble(interaction.depth=c(15), 
                              n.trees = 200,
                              shrinkage=c(0.1),
                              n.minobsinnode=40))
```
